---
title: "Template"
author: "Erik-Jan van Kesteren"
output:
  html_document:
    code_folding: hide
    includes:
      in_header: postheader.html
    keep_md: yes
    self_contained: no
    theme: journal
---

```{r setup, include=FALSE, echo=FALSE,render=FALSE,warning=FALSE,background=TRUE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev = 'svg')
```

<style type="text/css">
.table {

    margin-left: auto;
    margin-right: auto;
    display: block;

}
blockquote {

    font-size: 1em;
  
}
</style>

###[Back to index](../index.html)



## Introduction
Here is the intro and where I got the inspiration. It leads to this svg in the img folder of the root directory:

<img src="template_files/Isometric-blim.svg" width="40%" style="display:block; margin-left:auto; margin-right:auto"></img>

## The post
Here is a png dev image, for if there are a lot of points in the image (svg becomes slow in this case).

```{r png, tidy=TRUE, dpi = 400, dev='png'}

plot(rnorm(1000))

```

With few points, it's better to use the standard svg.
```{r svg, tidy=TRUE}

barplot(c(2,4,6,5,1), col="#00008B", border = "#00008B")

```

###[Back to index](../index.html)
